---
title: "HTTP Requests"
output: html_document
date: "2024-06-11"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("C:\\Users\\penza\\Documents\\Ed\\IBM Data Analytics with Excel and R\\5 Intro to R Programming")
```

HTTP Methods| Description
------------|-------------
GET         | Retrieves data from the server
POST        | Submits data to server
PUT         | Updates data already on server
DELETE      | Deletes data from server

[HTTP response status codes](https://developer.mozilla.org/en-US/docs/Web/HTTP/Status?utm_medium=Exinfluencer&utm_source=Exinfluencer&utm_content=000026UJ&utm_term=10006555&utm_id=NA-SkillsNetwork-Channel-SkillsNetworkCoursesIBMDeveloperSkillsNetworkRP0101ENCoursera889-2022-01-01)

```{r}
library(httr)
```

You can make a GET request via the method get to www.ibm.com:

```{r}
url<-'https://www.ibm.com/'
response<-GET(url)
response
```


We have the response object response , this has information about the response, like the status of the request. We can view the status code using the attribute status

```{r}
response$status
```
You can also check the headers of the response
```{r}
response_headers <- headers(response)
response_headers
```

We can obtain the date the request was sent using the key Date

```{r}
response_headers['date']
```

Content-Type indicates the type of data:

```{r}
response_headers['content-type']
```
To obtain the original request, you can view it via response object:

```{r}
response$request$headers
```

Coding Exercise: in the code cell below, find the content-length attribute in the response header

```{r}
response_headers['content-length']
```
Now, let's get the content of HTTP response
```{r}
content(response)
```
which is the IBM home page (in fact, HTML page which you will learn later in this course)

You can load other types of data for non-text requests like images, consider the URL of the following image:

```{r}
image_url<-'https://gitlab.com/ibm/skills-network/courses/placeholder101/-/raw/master/labs/module%201/images/IDSNlogo.png'
```

We can make a get request
```{r}
image_response<-GET(image_url)
```

We can look at the response header

```{r}
image_headers <- headers(image_response)
```

We can see the "Content-Type", which is an image
```{r}
image_headers['content-type']
```
An image is a response object that contains the image as a bytes-like object. As a result, we must save it using a file object. First, we specify the file path and name

```{r}
image <- content(image_response, "raw")
writeBin(image, "logo.png")
```
Then you should be able to find the log.png at the file explorer on the left

Coding Exercise: in the code cell below, find another image url and use above code to request and download the image

Get Request with URL Parameters 
You can also add URL parameters to HTTP GET request to filter resources. For example, instead of return all users from an API, I only want to get the user with id 1. To do so, I can add a URL parameter like userid = 1 in my GET request.

Let's see an GET example with URL parameters:

Suppose we have a simple GET API with base URL for http://httpbin.org/

```{r}
url_get <- 'http://httpbin.org/get'
```

and we want to add some URL parameters to above GET API. To do so, we simply create a named list with parameter names and values:

```{r}
query_params <- list(name = "Yan", ID = "123")
```
Then passing the list query_params to the query argument of the  GET() function.

It basically tells the GET API I only want to get resources with name equals Yan and id equals 123.

OK, let's make the GET request to 'http://httpbin.org/get' with the two arameters

```{r}
response <- GET(url_get, query=query_params)
```
We can print out the updated URL and see the attached URL parameters.

```{r}
response$request$url
```
After the base URL http://httpbin.org/get, you can see the URL parameters name=Yan&ID=123 are seperated by ?

The attribute args of the response had the name and values:
```{r}
content(response)$args
```

Post Requests 
Like a GET request a POST is used to send data to a server in a request body. In order to send the Post Request in Python in the URL we change the route to POST:
```{r}
url_post <- 'http://httpbin.org/post'
```
This endpoint will expect data as a file or as a form, a from is convenient way to configure an HTTP request to send data to a server.

To make a POST request we use the POST() function, the list body is passed to the parameter  body :

```{r}
body<- list(course_name='Introduction to R', instructor='Yan')
response<-POST('http://httpbin.org/post', body = body)
response
```
We can see POST request has a body stored in fields attribute

```{r}
response$request$fields
```

